Simple Models of Metal-Line Absorption and Emission from Cool Gas Outflows

### 2011-02-16 Winds/papers/Theory/First/Preprint/windsimple_prpt.tex
jxp
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5644 ac11df5e-a788-c23b-a8e1-c758c323924d

\documentclass[12pt,preprint]{aastex}
\usepackage{natbib,amsmath}

\special{papersize=8.5in,11in}
\begin{document}

@@

\begin{abstract}
We analyze the absorption and emission-line profiles produced by a set
of simple, cool gas wind models motivated by galactic-scale outflow
observations.  
We implement monte carlo radiative
transfer techniques that track the propagation of scattered and
fluorescent photons to generate 1D spectra and 2D spectral images.
{\bf [Should we make the codes publicly available?  Upon request?]}
%We examine outflows with a range of power-law density and velocity
%profiles and a radiative-driven wind model.  
We focus on the \mgiid\ doublet and \ion{Fe}{2} UV1
multiplet at $\lambda \approx 2600$\AA, but the results are applicable
to other transitions that trace outflows (e.g.\ \ion{Na}{1},
\ion{H}{1} \lya, \ion{Si}{2}).   
By design, the resonance transitions show blue-shifted
absorption but one also predicts strong resonance and fine-structure line-emission at roughly
the systemic velocity.  This line-emission `fills-in' the absorption
reducing the equivalent width by up to $50\%$, shifting the centroid
by tens of \kms, and severely
reducing the effective
opacity near systemic.  Analysis of cool gas outflows that ignores this
line-emission may incorrectly infer that the gas is partially
covered, measure a significantly lower peak optical depth, and/or
conclude that gas at systemic velocity is absent   
(e.g.\ an interstellar or slowly infalling component). 
Because the \ion{Fe}{2} lines
are connected by optically-thin transitions to
fine-structure levels, their profiles
more closely reproduce the intrinsic opacity
of the wind.   Together these results naturally explain the
absorption and emission-line characteristics observed for star-forming
galaxies at $z<1$.
We also study a scenario purported to describe the outflows
of $z \sim 3$ Lyman break galaxies and argue this model fails to
reproduce the observations when radiative transfer of scattered photons
is properly considered.   
%Strong line-emission from metal-line
%transitions is a generic prediction of wind models, even in the
%presence of dust and anisotropic configurations.  
Although line-emission complicates the analysis of absorption-line
profiles, the 
surface brightness profiles 
offers a unique means of assessing the morphology and size
of galactic-scale winds.  Furthermore, the kinematics and line-ratios
offer powerful diagnostics of outflows,  motivating deep,
spatially-extended spectroscopic observations.

@@

Nearly all gaseous objects that shine are also
observed to generate gaseous flows.  This includes the jets of protostars, the
stellar winds of massive O and B stars, the gentle Solar wind
of our Sun, the associated absorption of bright quasars, and the
spectacular jets of radio-loud AGN.   These gaseous outflows 
%%KHRR
regulate the metal and dust content and distribution within the 
objects and its surroundings, 
%%
moderate the accretion of new material, and %onto the object, 
inject energy and momentum into gas on large scales. %, and ...
Developing a comprehensive model for these flows is critical to
understanding the evolution of the source and its impact on the
surrounding environment.

Starburst galaxies, whose luminosity is dominated by \ion{H}{2} regions
and massive stars, are also observed to drive gaseous outflows.  These
flows are generally expected (and sometimes observed) to have multiple
phases, for example a hot and diffuse phase traced by X-ray emission
together with a cool, denser phase traced by H$\alpha$ emission 
%% KHRR -- added refs
\citep[e.g.][]{ham90,martin99,shc+04,km10}. 
Several spectacular
examples in the local universe demonstrate that flows can extend to
up to $\sim 10$ kpc
%tens of kpc 
from the galaxy \citep{lhw99,wsg08,vsr+03} carrying significant speed
to escape from the gravitational potential well of the galaxy's dark
matter halo \citep[e.g.][]{sh09}.

%Beyond $z \sim 0$, 
Galactic outflows are also
revealed by UV and optical absorption lines, e.g.\ \ion{Na}{1},
\ion{Mg}{2}, \ion{Si}{2} and \ion{C}{4} transitions.  With the galaxy
as a backlight, one observes gas that is predominantly
blue-shifted which indicates a flow toward
Earth and away from the galaxy.  These transitions are sensitive to
the cool (\ion{Mg}{2}; $T \sim 10^4$K) and warm (\ion{C}{4}; $T \sim
10^5$K) phases of the flow.  
The incidence of cool gas outflows is
nearly universal in 
%% KHRR added:
vigorously
%%
star-forming galaxies;  this includes systems at $z \sim 0$
which exhibit \ion{Na}{1} and \ion{Mg}{2} absorption
\citep{rvs05a,martin05,smn+09,mb09,cth+10}, 
$z \sim 1$ star-forming galaxies traced by
\ion{Fe}{2} and \ion{Mg}{2} transitions \citep{wcp+09,rwk+10}, and
$z>2$ Lyman break galaxies (LBGs) that show blue-shifted \ion{Si}{2},
\ion{C}{2}, and \ion{O}{1} transitions \citep{sgp+96,lkg+97,shapley03}.

@@

Recent studies of $z \lesssim 1$ star-forming galaxies have revealed that
the cool outflowing gas often exhibits significant resonant-line emission (e.g.\
\ion{Mg}{2}, \ion{Na}{1}) in
tandem with the nearly ubiquitous blue-shifted absorption
\citep{wcp+09,rwk+10,cth+10}.  The resultant spectra resemble the P-Cygni
profile characteristic of stellar winds.
This phenomenon was first reported by \citet{phillips93}, who observed 
blue-shifted absorption and red-shifted emission for the \ion{Na}{1}
transition in the spectrum of the local starburst galaxy NGC 1808.
More recently, \cite{wcp+09}, who
studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
\ion{Mg}{2} emission in a small subset of the individual galaxy
spectra of their large sample.  These were excluded from the full
analysis on concerns that the emission was related to AGN activity.
The stacked spectra of the remaining galaxies, however, also indicated
\ion{Mg}{2} emission, both directly and when the authors modeled and
`removed' the $v \approx 0\mkms$ absorption component.  The
authors suggested the emission could be related to back-scattered
light in the wind, but presumed that it was related to weak
AGN activity.   Very similar \ion{Mg}{2} emission was observed by
\cite{rwk+10} who repeated the analysis of \cite{wcp+09} on a set
of lower redshift galaxies.

Bright \ion{Mg}{2} line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite[][see also
Rubin et al\ 2011, in prep.]{rubin+10c}.
In their analysis of a single galaxy spectrum, \cite{rubin+10c}
further demonstrated that the \ion{Mg}{2} line emission is spatially
extended, and used the size of the emission to infer that the wind 
extends to at least 7\,kpc from the galaxy.
These authors additionally detected line
emission from non-resonant \ion{Fe}{2}$^*$ transitions, and attribute
the emission to fluorescence powered by \ion{Fe}{2} resonant
absorption.  In other words, these photons are re-emitted by the wind
into our sightline, and are analogous to the emitted photons in a P-Cygni profile.
%Their interpretation is that the emission is the result of photons
%scattered off the wind into our sightline, analogous to P-Cygni.
%\cite{rubin+10c} also observed spatially extended \ion{Mg}{2} emission
%which they use to estimate the size of the outflow.  They infer that
%the wind extends to at least 7\,kpc from the galaxy.  
Line-emission is
also observed for $z \sim 3$ LBGs in the resonant \lya\ transition
and non-resonant \ion{Si}{2}$^*$ transitions\footnote{See also
  \cite{france10} for a new example at $z \sim 0$.} \citep{prs+02,shapley03}.
We emphasize that this emission arises from a completely different
physical process than those generating X-ray and H$\alpha$ emission
(e.g., shocks), and presumably probes both the cool gas at the base of
the wind and the outskirts of the flow (i.e., wherever a given
transition is optically thick).
A comprehensive analysis of the %emission lines
scattered and fluorescent emission
related to galactic-scale outflows
(e.g.\ via deep integral-field-unit [IFU] observations,
\citealt{ssb+05,wsg08}) 
may offer unique
diagnostics on the spatial extent, morphology, and density of the outflow 
from distant galaxies,
eventually
setting tighter constraints on the energetics of the flow.

@@

@@

In a resonant line interaction, a photon excites an atom from the ground state to an excited level.  The end result is either scattering (i.e., de-excitation back to the ground state) or fluorescence (de-excitation to another excited  level).  The probability that the atom de-excites to lower level $x$ is  $p_{x} = A_{ux} / \sum_i A_{ui}$, where the $A_{ui}$ are the Einstein spontaneous emission coefficients and the sum runs over all levels accessible from the upper state $u$.  
For each interaction event, a random number was drawn to sample the final state from this probability distribution.    If the result was a scattering, the process was assumed to be coherent in the comoving frame.  If the result was fluorescence, the packet was  reemitted at the line center wavelength of the new transition.  Natural line broadening in fluorescence was ignored given the high velocity gradients in our models.  In all cases, the angular redistribution was  assumed to be isotropic.

%Spatially averaged spectra of the models were constructed by collecting all packets which escaped in a given (observer-frame) wavelength bin.   
To generate multi-dimensional images and spectra of the system, an escape probability method was used.  At the initiation of a packet, and at every subsequent interaction event, we calculated the probability $P_x = e^{-\tau_{ux}} p_{ux}$,  that the packet de-excited into state $x$ and escaped the domain in some pre-specified direction. Here
$\tau_{ux}$ is the optical depth to infinity, which was  constructed  by integration along the path to escape, taking into account the relevant Doppler shifts and possible absorption by lines or dust.  The contribution of the packet to the final spectrum was than added in for every possible final state, each shifted to the proper observer-frame wavelength and weighted by the probability $P_x$.

To validate the monte carlo code, we calculated a series of test problems consisting of a point source embedded in a spherical, homogenous medium of varying resonant line optical depth.  The resulting resonant line profiles  \citep[see][]{Kasen_lyman} were
found to be in good agreement with the standard analytic solutions
\citep{Harrington_1973,Neufeld_1990,Dijkstra_2006}.  As a further
test, we calculated line profiles for the case of an extended
spherical source embedded in a homologously expanding wind with large
velocity gradient.  These model spectra
(Figure~\ref{fig:oneline_test}) are in good agreement with ones
determined by direct numerical integration of the radiation transport
equation under the Sobolev approximation \citep[e.g.,][]{Jeffery_Branch}.

@@

A photon propagating through a differentially expanding medium
interacts with a line only when
it's comoving frame wavelength is Doppler shifted into near resonance with the line center rest wavelength. 
For a wind with a steep velocity law  (i.e.\ a large gradient $dv/dr$)
and/or a narrow intrinsic profile (i.e.\ a small Doppler parameter), the 
spatial extent of the region of resonance
may be much smaller than the length scale of the
wind itself.  The interaction can then 
be considered to occur at  a point.
In this case, \citet{sobolev60} introduced a formalism that gives the line optical depth $\tau(r)$ at  
a given point in terms of the density and
velocity gradient of the flow at that radius.  For a wind in homologous expansion,
this optical depth is independent of the direction of propagation and
is given by
\begin{equation}
\tau_S(r) = K_\ell   \lambda^0_\ell | dv/dr |^{-1} \cmma
\label{eqn:Sobolev}
\end{equation}
where 
\begin{equation}
K_\ell = \frac{\pi e^2}{m_e c} f_\ell n_\ell(r)
\end{equation}
is the integrated line opacity, $\lambda^0_\ell$ the rest center wavelength,
$f_\ell$ the oscillator strength and
$n_\ell$ the density in the lower level of the transition.   We have
neglected corrections for stimulated emission.   This optical
depth applies to a photon with wavelength $\lambda_S
\equiv \lambda_\ell (1- v(r)~\hat{r}  \cdot \hat{d} /c)$ where $\hat{d}$ is the 
direction of propagation.  The probability that such a photon
is scattered/absorbed at the point of resonance is simply $1 - \exp[-\tau_S(r)]$.
We find that the Sobolev approximation applies for nearly all of the
models presented in this paper, and therefore provides a convenient
approach to estimating the optical depth.

@@

The fiducial wind is isotropic, dust-free, and extends from an inner wind
radius $r_{\rm inner}$ to an outer wind radius $r_{\rm outer}$.  
It follows a (mass-flux conserving) density law,

@@

@@

For the spatially integrated (i.e., 1D) spectra, the actual dimensions and density of the wind are
unimportant provided they scale together to give nearly the same
optical depth. Therefore, one may consider the choices for
$r_{\rm inner}, r_{\rm outer}, n_{\rm H}^0$ as largely arbitrary.
Nevertheless, we adopted values for this fiducial model with
some astrophysical motivation,  e.g., values that correspond to
galactic dimensions and a normalization that gives $\tau_{2796}^{\rm
  max} \sim 10$.

Using the methodology described in $\S$~\ref{sec:method}, we
propagated photons from the source and through the outflow to an
`observer' at $r \gg r_{\rm outer}$ that views the entire wind+source
complex.  Figure~\ref{fig:fiducial_1d} presents the 1D spectrum
that this observer would record, with the unattenuated flux
normalized to unit value.   The \ion{Mg}{2} doublet
shows the canonical `P-Cygni profile' that characterizes a continuum
source embedded within an outflow.  Strong absorption is evident at
$\delta v  \equiv (\lambda/\lambda_0 - 1) < -50 \mkms$ in both transitions (equivalent widths, $W^{2796} =
3.0$\AA\ and $W^{2803} = 1.3$\AA) and each shows emission at
positive velocities.  For an isotropic and dust-free model, the
total equivalent width of the doublet must be zero,
i.e.\ every photon
absorbed eventually escapes the system, typically at lower
energy.  The wind simply shuffles the photons in frequency space.
A simple summation of the absorption and emission equivalent widths
(Table~\ref{tab:fiducial_EW}) confirms this expectation.

Focusing further on the \ion{Mg}{2} absorption, one notes that the profiles lie
well above zero intensity and have similar depth even though their $f\lambda$
values differ by a factor of two.  In standard absorption-line
analysis, this is 
the tell-tale signature of a `cloud' that has a high optical depth (i.e.\
saturated) which only partially covers the emitting source
\citep[e.g.][]{rvs05a,hkp+10}.  Our fiducial wind model, however, 
{\it entirely covers the source}; the apparent partial covering must
be related to a different effect.
Figure~\ref{fig:noemiss} further emphasizes this point by comparing the 
absorption profiles from Figure~\ref{fig:fiducial_1d} against an
artificial model where no absorbed photons are 
re-emitted.   As expected from the
$\tau_{2796}$ profile (Figure~\ref{fig:fiducial_nvt}), this
`intrinsic' model
produces a strong \mgiid\ doublet that absorbs all photons at
$\delta v \approx -100 \mkms$, i.e.\ $I_{2796}^{\rm min} = \exp(-\tau^{\rm
  max}_{2796}) \approx 0$.
The true model, in contrast, has been `filled in' at $\delta v \approx -100
\mkms$ by photons scattered in the wind.  An
absorption-line analysis that ignores these effects
would (i) systematically underestimate the true optical
depth and/or (ii) falsely conclude that the wind partially covers the
source.  We will find that these are
generic results, even for wind models that include
dust and are not fully isotropic.

>>>>

\documentclass[]{emulateapj}
\usepackage{natbib,amsmath}

\special{papersize=8.5in,11in}
\shorttitle{Simple Models of Cool Gas Outflows} 
\shortauthors{Prochaska, Kasen, \& Rubin}

\begin{document}

@@

\begin{abstract}
We analyze the absorption and emission-line profiles produced by a set
of simple, cool gas wind models motivated by galactic-scale outflow
observations.  
We implement monte carlo radiative
transfer techniques that track the propagation of scattered and
fluorescent photons to generate 1D spectra and 2D spectral images.
%{\bf [Should we make the codes publicly available?  Upon request?]}
%We examine outflows with a range of power-law density and velocity
%profiles and a radiative-driven wind model.  
We focus on the \mgiid\ doublet and \ion{Fe}{2} UV1
multiplet at $\lambda \approx 2600$\AA, but the results are applicable
to other transitions that trace outflows (e.g.\ \ion{Na}{1},
\ion{H}{1} \lya, \ion{Si}{2}).   
By design, the resonance transitions show blue-shifted
absorption but one also predicts strong resonance and fine-structure line-emission at roughly
the systemic velocity.  This line-emission `fills-in' the absorption
reducing the equivalent width by up to $50\%$, shift the
absorption-lin centroid by tens of \kms, and 
reduce the effective
opacity near systemic.  Analysis of cool gas outflows that ignores this
line-emission may incorrectly infer that the gas is partially
covered, measure a significantly lower peak optical depth, and/or
conclude that gas at systemic velocity is absent   
(e.g.\ an interstellar or slowly infalling component). 
Because the \ion{Fe}{2} lines
are connected by optically-thin transitions to
fine-structure levels, their profiles
more closely reproduce the intrinsic opacity
of the wind.   Together these results naturally explain the
absorption and emission-line characteristics observed for star-forming
galaxies at $z<1$.
We also study a scenario promoted to describe the outflows
of $z \sim 3$ Lyman break galaxies and 
find prfiles inconsistent with the observations due to scattered
photon emission.
%argue this model fails to
%reproduce the observations when radiative transfer of scattered photons
%is properly considered.   
%Strong line-emission from metal-line
%transitions is a generic prediction of wind models, even in the
%presence of dust and anisotropic configurations.  
Although line-emission complicates the analysis of absorption-line
profiles, the 
surface brightness profiles 
offer a unique means of assessing the morphology and size
of galactic-scale winds.  Furthermore, the kinematics and line-ratios
offer powerful diagnostics of outflows,  motivating deep,
spatially-extended spectroscopic observations.

@@

Nearly all gaseous objects that shine are also
observed to generate gaseous flows.  This includes the jets of protostars, the
stellar winds of massive O and B stars, the gentle Solar wind
of our Sun, the associated absorption of bright quasars, and the
spectacular jets of radio-loud AGN.   These gaseous outflows 
regulate the metal and dust content and distribution within the 
objects and their surroundings, 
moderate the accretion of new material, and %onto the object, 
inject energy and momentum into gas on large scales. %, and ...
Developing a comprehensive model for these flows is critical to
understanding the evolution of the source and its impact on the
surrounding environment.

Starburst galaxies, whose luminosity is dominated by \ion{H}{2} regions
and massive stars, are also observed to drive gaseous outflows.  These
flows are generally expected (and sometimes observed) to have multiple
phases, for example a hot and diffuse phase traced by X-ray emission
together with a cool, denser phase traced by H$\alpha$ emission 
\citep[e.g.][]{ham90,martin99,shc+04,km10}. 
Several spectacular
examples in the local universe demonstrate that flows can extend to
up to $\sim 10$ kpc
from the galaxy \citep{lhw99,vsr+03,wsg08} carrying significant speed
to escape from the gravitational potential well of the galaxy's dark
matter halo \citep[e.g.][]{sh09}.

Galactic outflows are also
revealed by UV and optical absorption lines, e.g.\ \ion{Na}{1},
\ion{Mg}{2}, \ion{Si}{2} and \ion{C}{4} transitions.  With the galaxy
as a backlight, one observes gas that is predominantly
blue-shifted which indicates a flow toward
Earth and away from the galaxy.  These transitions are sensitive to
the cool (\ion{Mg}{2}; $T \sim 10^4$K) and warm (\ion{C}{4}; $T \sim
10^5$K) phases of the flow.  
The incidence of cool gas outflows is
nearly universal in vigorously
star-forming galaxies;  this includes systems at low $z$
which exhibit \ion{Na}{1} and \ion{Mg}{2} absorption
\citep{rvs05a,martin05,smn+09,mb09,cth+10}, 
$z \sim 1$ star-forming galaxies with winds traced by
\ion{Fe}{2} and \ion{Mg}{2} transitions \citep{wcp+09,rwk+10}, and
$z>2$ Lyman break galaxies (LBGs) that show blue-shifted \ion{Si}{2},
\ion{C}{2}, and \ion{O}{1} transitions \citep{sgp+96,lkg+97,shapley03}.

@@

Recent studies of $z \lesssim 1$ star-forming galaxies have revealed that
the cool outflowing gas often exhibits significant resonant-line emission (e.g.\
\ion{Mg}{2}, \ion{Na}{1}) in
tandem with the nearly ubiquitous blue-shifted absorption
\citep{wcp+09,mb09,rwk+10,cth+10}.  The resultant spectra resemble the P-Cygni
profile characteristic of stellar winds.
This phenomenon was first reported by \citet{phillips93}, who observed 
blue-shifted absorption and red-shifted emission for the \ion{Na}{1}
transition in the spectrum of the local starburst galaxy NGC 1808.
More recently, \cite{wcp+09}, who
studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
\ion{Mg}{2} emission in a small subset of the individual galaxy
spectra of their large sample.  These were excluded from the full
analysis on concerns that the emission was related to AGN activity.
The stacked spectra of the remaining galaxies, however, also indicated
\ion{Mg}{2} emission, both directly and when the authors modeled and
`removed' the $v \approx 0\mkms$ absorption component.  The
authors suggested the emission could be related to back-scattered
light in the wind, but presumed that it was related to weak
AGN activity.   Very similar \ion{Mg}{2} emission was observed by
\cite{rwk+10} who repeated the analysis of \cite{wcp+09} on a set
of lower redshift galaxies.

Bright \ion{Mg}{2} line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite[][see also
Rubin et al\ 2011, in prep.]{rubin+10c}.
In their analysis of a single galaxy spectrum, \cite{rubin+10c}
further demonstrated that the \ion{Mg}{2} line emission is spatially
extended, and used the size of the emission to infer that the wind 
extends to at least 7\,kpc from the galaxy.
These authors additionally detected line
emission from non-resonant \ion{Fe}{2}$^*$ transitions, and attributed
the emission to fluorescence powered by \ion{Fe}{2} resonant
absorption.  In other words, these photons are re-emitted by the wind
into our sightline, and are analogous to the emitted photons in a P-Cygni profile.
%Their interpretation is that the emission is the result of photons
%scattered off the wind into our sightline, analogous to P-Cygni.
%\cite{rubin+10c} also observed spatially extended \ion{Mg}{2} emission
%which they use to estimate the size of the outflow.  They infer that
%the wind extends to at least 7\,kpc from the galaxy.  
Line-emission that may be related to outflows is
also observed for $z \sim 3$ LBGs in the resonant \lya\ transition
and non-resonant \ion{Si}{2}$^*$ transitions\footnote{See also
  \cite{france10} for a new example at $z \sim 0$.} \citep{prs+02,shapley03}.
This emission likely arises from a completely different
physical process than those generating X-ray and H$\alpha$ emission
(e.g., shocks), and presumably probes both the cool gas at the base of
the wind and the outskirts of the flow (i.e., wherever a given
transition is optically thick).
A comprehensive analysis of the %emission lines
scattered and fluorescent emission
related to galactic-scale outflows
(e.g.\ via deep integral-field-unit [IFU] observations,
\citealt{ssb+05,wsg08}) 
may offer unique
diagnostics on the spatial extent, morphology, and density of the outflow 
from distant galaxies,
eventually
setting tighter constraints on the energetics of the flow.

@@

\begin{deluxetable}{lcccccc}
\tabletypesize{\footnotesize}
\tablecolumns{11}
\tablecaption{\ion{Mg}{2} and \ion{Fe}{2} Transitions Considered \label{tab:atomic}}
\tablewidth{0pt}
\tablehead{\colhead{} & \colhead{$\rm E_{high}$} & \colhead{$\rm E_{low}$} & \colhead{$J_{\rm high}$} & \colhead{$J_{\rm low}$} & \colhead{$\lambda$} & \colhead{$A$} \\
 & \colhead{($\rm cm^{-1}$)} & \colhead{($\rm cm^{-1}$)} &&& \colhead{(\AA)} & \colhead{($\rm s^{-1}$)} } 
\startdata
\ion{Fe}{2} UV1 & 38458.98 &     0.00 &   9/2 & 9/2 & 2600.173 &
2.36E+08  \\
           & 38458.98 &   384.79 &   9/2 & 7/2 & 2626.451 & 3.41E+07 \\
           & 38660.04 &     0.00 &   7/2 & 9/2 & 2586.650 & 8.61E+07 \\
           & 38660.04 &   384.79 &   7/2 & 7/2 & 2612.654 & 1.23E+08 \\
           & 38660.04 &   667.68 &   7/2 & 5/2 & 2632.108 & 6.21E+07 \\
           & 38858.96 &   667.68 &   5/2 & 5/2 & 2618.399 & 4.91E+07 \\
           & 38858.96 &   862.62 &   5/2 & 3/2 & 2631.832 & 8.39E+07 \\
\tableline \\ [-1.5ex]
\ion{Mg}{2}& 35760.89 &     0.00 &   3/2 &   0 & 2796.351 & 2.63E+08\\
           & 35669.34 &     0.00 &   1/2 &   0 & 2803.528 & 2.60E+08\\
\enddata
\tablecomments{Atomic data from \citet{morton03}.}
\end{deluxetable}

\begin{figure}
%\epsscale{0.95}
%\plotone{../Figures/energy_levels.ps}
\includegraphics[width=3.5in]{f1.eps}
\caption{
Energy level diagrams for the \mgiid\ doublet and the UV1
multiplet of \ion{Fe}{2} transitions   
(based on Figure~7 from \cite{hmt+99}).
Each transition shown is
labeled by its rest wavelength (\AA) and Einstein A-coefficient
(s$^{-1}$). Black upward arrows
indicate the resonance-line transitions, i.e.\ those connected to the ground
state.  The 2p$^6$3p configuration of Mg$^+$ is split into
two energy levels that give rise to the \mgiid\ doublet.  
Both the 3d$^6$4s ground state and 3d$^6$4p upper level of Fe$^+$
exhibit fine-structure splitting that gives rise to a series of
electric-dipole transitions. 
The downward (green) arrows show the \feiis\ transitions that are connected to the
resonance-line transitions (i.e.\ they share the same upper energy
levels).  We also show a pair of transitions (\ion{Fe}{2}$^* \lambda\lambda
2618,2631$; red and dashed lines) that arise from higher levels in the \zconfig\
configuration.  These transitions have not yet been observed in
galactic-scale outflows and are not considered in our analysis.
}
\label{fig:energy}
\end{figure}

@@

In a resonant line interaction, a photon excites an atom from the
ground state to an excited level.  The end result is either scattering
(i.e., de-excitation back to the ground state) or fluorescence
(de-excitation to another excited  level).  The probability that the
atom de-excites to lower level $x$ is  $p_{ux} = A_{ux} / \sum_i A_{ui}$, where the $A_{ui}$ are the Einstein spontaneous emission coefficients and the sum runs over all levels accessible from the upper state $u$.  
For each interaction event, a random number was drawn to sample the
final state from this probability distribution.    If the result was a
scattering, the process was assumed to be coherent in the comoving
frame.  If the result was fluorescence, the packet was  reemitted at
the line center wavelength of the new transition.  Natural line
broadening in fluorescence was ignored given the high velocity
gradients in our models.  In all cases, the angular redistribution was
assumed to be isotropic.

%Spatially averaged spectra of the models were constructed by collecting all packets which escaped in a given (observer-frame) wavelength bin.   
To generate multi-dimensional images and spectra of the system, an escape probability method was used.  At the initiation of a packet, and at every subsequent interaction event, we calculated the probability $P_x = e^{-\tau_{ux}} p_{ux}$,  that the packet de-excited into state $x$ and escaped the domain in some pre-specified direction. Here
$\tau_{ux}$ is the optical depth to infinity, which was  constructed
by integration along the path to escape, taking into account the
relevant Doppler shifts and possible absorption by lines or dust.  The
contribution of the packet to the final spectrum was then added in for
every possible final state, each shifted to the proper observer-frame
wavelength and weighted by the probability $P_x$.

\begin{figure}
%\epsscale{0.8}
%\plotone{../Figures/oneline_test.eps}
\includegraphics[width=3.5in]{f2.eps}
\caption{
Line profiles of a test problem calculated using the monte carlo code (solid) and by direct numerical integration of the transport equation under the Sobolev approximation (dashed lines).  The configuration was  spherical and consisted of a homologously expanding wind  with density profile $n \propto r^{-2}$  lying above an opaque emitting surface at velocity coordinate 100~km~s$^{-1}$.  
The wind extended to 2,000 km~s$^{-1}$ and the Sobolev optical depth
had values of  $\tau = 1$ (black lines), $\tau = 10$ (red lines), and
$\tau = 100$ (blue lines).    Discrepancies in the $\tau = 100$ case
are likely due to the breakdown of the Sobolev approximation
($\S$~\ref{sec:Sobolev}) at high optical depth. 
}
\label{fig:oneline_test}
\end{figure}

To validate the monte carlo code, we calculated a series of test
problems consisting of a point source embedded in a spherical,
homogenous medium of varying resonant line optical depth.  The
resulting resonant line profiles  \citep[see][]{Kasen_lyman} were 
found to be in good agreement with the standard analytic solutions
\citep{Harrington_1973,Neufeld_1990,Dijkstra_2006}.  As a further
test, we calculated line profiles for the case of an extended
spherical source embedded in a homologously expanding wind with large
velocity gradient.  These model spectra
(Figure~\ref{fig:oneline_test}) are in good agreement with ones
determined by direct numerical integration of the radiation transport
equation under the Sobolev approximation
\citep[$\S$~\ref{sec:Sobolev}; e.g.,][]{Jeffery_Branch}.

@@

A photon propagating through a differentially expanding medium
interacts with a line only when
its comoving frame wavelength is Doppler-shifted into near resonance with the line center rest wavelength. 
For a wind with a steep velocity law  (i.e.\ a large gradient $dv/dr$)
and/or a narrow intrinsic profile (i.e.\ a small Doppler parameter), the 
spatial extent of the region of resonance
may be much smaller than the length scale of the
wind itself.  The interaction can then 
be considered to occur at  a point.
In this case, \citet{sobolev60} introduced a formalism that gives the line optical depth $\tau(r)$ at  
a given point in terms of the density and
velocity gradient of the flow at that radius.  For a wind in homologous expansion,
this optical depth is independent of the direction of propagation and
is given by
\begin{equation}
\tau_S(r) = K_\ell   \lambda^0_\ell | dv/dr |^{-1} \cmma
\label{eqn:Sobolev}
\end{equation}
where 
\begin{equation}
K_\ell = \frac{\pi e^2}{m_e c} f_\ell n_\ell(r)
\end{equation}
is the integrated line opacity, $\lambda^0_\ell$ the line-center rest wavelength,
$f_\ell$ the oscillator strength and
$n_\ell$ the density in the lower level of the transition.   We have
neglected corrections for stimulated emission.   This optical
depth applies to a photon with wavelength $\lambda_S
\equiv \lambda_\ell (1- v(r)~\hat{r}  \cdot \hat{d} /c)$ where $\hat{d}$ is the 
direction of propagation.  The probability that such a photon
is scattered/absorbed at the point of resonance is simply $1 - \exp[-\tau_S(r)]$.
We find that the Sobolev approximation applies for nearly all of the
models presented in this paper, and therefore provides a convenient
approach to estimating the optical depth.

@@

The fiducial wind is isotropic, dust-free, and extends from an inner wind
radius $r_{\rm inner}$ to an outer wind radius $r_{\rm outer}$.  
%It follows a (mass-flux conserving) density law,
It follows a density law,

@@

\begin{deluxetable}{ccl}
\tablewidth{0pc}
\tablecaption{Wind Parameters: Fiducial Model\label{tab:fiducial}}
\tabletypesize{\footnotesize}
\tablehead{\colhead{Property} & \colhead{Parameter} & \colhead{Value} } 
\startdata
Density law  & $n(r)$ & $\propto r^{-2}$ \\
Velocity law  & $v_r$ & $ \propto r$ \\
Inner Radius & $r_{\rm inner}$ & 1\,kpc \\
Outer Radius & $r_{\rm outer}$ & 20\,kpc \\
Source size  & $r_{\rm source}$ & 0.5\,kpc \\
Density Normalization & $n^0_{\rm H}$ & $0.1\cm{-3}$ at $r_{\rm inner}$ \\
Velocity Normalization & $v_0$ & 1000\kms at $r_{\rm outer}$ \\
Turbulence   & $b_{\rm turb}$  & 15 \kms \\
Mg$^+$ Normalization & \nmg\ & $10^{-5.47} n_{\rm H}$ \\
Fe$^+$ Normalization & \nfe\ & \nmg/2 \\
\enddata
%\tablecomments{Unless specified otherwise, all quantities refer to the \sna=2 threshold.  The cosmology assumed has $\Omega_\Lambda = 0.7, \Omega_m = 0.3$, and $H_0 = 72 \mkms \rm Mpc^{-1}$.}
%\tablenotetext{a}{Total redshift survey path for the \sna=2 criterion.}

\end{deluxetable}

@@

For the spatially integrated (i.e., 1D) spectra, the actual dimensions and density of the wind are
unimportant provided they scale together to give nearly the same
optical depth. Therefore, one may consider the choices for
$r_{\rm inner}, r_{\rm outer}$, and $n_{\rm H}^0$ as largely arbitrary.
Nevertheless, we adopted values for this fiducial model with
some astrophysical motivation,  e.g., values that correspond to
galactic dimensions and a normalization that gives $\tau_{2796}^{\rm
  max} \sim 10$.

\begin{figure}
%\epsscale{0.8}
%\plotone{../Figures/fig_nvtau_vs_r.ps}
\includegraphics[height=3.5in,angle=90]{f3.eps}
\caption{
Density (dashed; red), radial velocity (dotted; blue), and
\mgiia\ optical depth profiles (solid and dotted; black) for the fiducial
wind model (see Table~\ref{tab:fiducial} for details).
The density and velocity laws are simple $r^{-2}$ and $r^1$
power-laws; these curves have been scaled for plotting
convenience.  
The optical depth profile was calculated two ways: (i) using the
Sobolev approximation (Equation~\ref{eqn:Sobolev}; solid black curve)
and (ii) summing
the opacity at small and discrete radial intervals in velocity space
and then converting to radius with the velocity law (doted black
curve).  These give
very similar results. 
The wind parameters were set to give an optically thick medium at the
inner radius ($r_{\rm inner} = 1$\,kpc) that becomes optically thin at
the outer radius ($r_{\rm outer} = 20$\,kpc). 
}
\label{fig:fiducial_nvt}
\end{figure}

\begin{deluxetable}{ccr}
\tablewidth{0pc}
\tablecaption{Equivalent Widths for the Fiducial Model\label{tab:fiducial_EW}}
\tabletypesize{\footnotesize}
\tablehead{\colhead{Transition} & \colhead{$v_{\rm int}^a$} & \colhead{$W_{\lambda}^b$} \\ 
& (\kms) & (\AA) }
\startdata
  MgII 2796  &[$-1030,-65$]& 2.96\\
&[$-65,311$]&$-1.75$\\
  MgII 2803  &[$-458,-57$]& 1.29\\
&[$-57,692$]&$-2.34$\\
  FeII 2586  &[$-365,-46$]& 0.59\\
&[$-46,186$]&$-0.10$\\
  FeII 2600  &[$-626,-78$]& 1.16\\
&[$-49,470$]&$-0.94$\\
  FeII* 2612 &[$-219,183$]&$-0.33$\\
  FeII* 2626 &[$-194,177$]&$-0.27$\\
  FeII* 2632 &[$-155,130$]&$-0.16$\\
\enddata
\tablenotetext{a}{Velocity interval over which the equivalent width is calculated.}
\tablenotetext{b}{Negative values indicate line-emission.}
\end{deluxetable}
